### 2014-06-30 patchy-selection-paper.tex
put in dispersal refs

A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. \citep{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
On the lava flow the dark MC1R allele is at 86\% frequency,
while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
These numbers give us a sense of a plausible range of characteristic lengths.
Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
we can fit formula \eqref{eqn:eqfreq} to these frequencies.
Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
We also need estimate of $s_m$. 
Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
$1/\sqrt{30}$ are reasonable values.
\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
which has dispersal distance of 0.26km.  Shall we ask Nachman?}

@@

\subsection{PETER'S VERSION}

This paper is an investigation into the basic, ill-formed question: 
What is the spatial resolution of adaptation?
In other words, what scale of environmental patchiness can a species develop and sustain local adaptations,
and over what spatial scale will evolution develop independent solutions to evolutionary problems?
The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
the dispersal distance divided by the square root of the selection against the new allele between patches,
and much more weakly on the selective benefit within the patches or (perhaps suprisingly) the population density.
Weaker selection against alleles will make sharing of adaptations between patches easier,
but will also make it harder to spot,
since the lucky, transiting alleles move slower and carry a shorter haplotype.

Observations of similar adaptations occuring within the same species in widely separated patches
is strongly suggestive of parallel, independent adaptations.
Such observations are very interesting 
as they tell us about the degree of constraint faced by evolution --
how many different methods are there typically available to a species
with which to solve an evolutionary problem?

It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
\plr{fisher probably said this too?}
and that spatial structure contributes to this \citep{ralph2010parallel}.
Naturally, then, local adaptation to well-separated patches of some environment is even more likely to occur in parallel;
we have studied the case where patches are separated by long distances.
This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
Intermediate cases where habitat types are distributed in some more complex geometry
-- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
remains to be studied.

To complement this theory, we need more empirical work,
in which the parameters (migration rate, sensitivity to environmental factors, etc)
are reasonably well-understood.
However, learning about these aspects is difficult in the field,
and it is still not understood how much of the intraspecific variation we observe
is due to adaptation to local conditions, or even how to tell.
Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?

@@

\paragraph{Dominance}
We have mostly swept the issue of dominance under the rug
by dealing with essentially haploid models,
and appealing to the fact that the dynamics we study occur where the mutation is rare,
and hence mostly present only in heterozygotes.
If, however, the mutation is recessive, then it is essentially neutral where rare,
and so could have a much easier time spreading between patches.
This is counteracted, however, by the increased difficulty with which the mutation would establish
once arriving in a new patch.
On the other hand, a dominant mutation would not differ greatly from the theory presented here.
In fact, MC1r is ?dominant? and agouti is ?recessive?
Other schemes could be obtained.

optimal $s_m$ for migration makes migration harder to spot

\subsection{GRAHAM'S VERSION}

We have developed a simple set of approximations to describe the rate
at which organisms in a patchy environment adapt through novel
mutations or migration from other nearby patches (eqn.
\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
We have also determined the characteristic genomic length scales over which
the various signals of hitchhiking should be visable
(eqn. \eqref{eqn:haplotype_length} and ).

These results show that the probability of parallel adaptation is
independent of both the population density and the beneficial
selection coefficient. This follows from the fact that these
parameters both increase the rate of adaptation via both migration and
mutation equally. Thus they are important for the time-scale over
which adaptation occurs but not the mode of adaptation. This
constrasts with models of convergent adaptation within panmictic
populations \citep{} and geographically spread populations adapting to
homogeneous populations \citep{RalphCoop}, where the number of
adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
coefficient is absent or weak, due to the fact that selection both
aids establiishment of new alleles and the spread of existing
alleles.

The primary determinant of the probability of parallel adaptation is
$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
allele frequency decay off the patch. We find that patches that are close together on
this scale will adapt via migration, but there will be a sharp
transition to convergent adaptation as we move to patches that are further apart.

In coming to this conclusion, we have ignored the possibility of very long distance migration,
instead focusing local dispersal (as parameterized by a Gaussian
dispersal model). However, dispersal distributions can be very heavy
tailed, with a small fraction of individuals moving very long distances indeed, in
addition over long time-scales very rare chance events (mice sucked up in
hurricanes etc) could play a role in spreading migrant alleles. Such
tail events could greatly increase the probability of parallel
adaptation above that predicted by our model. Furthermore, if
adaptive alleles do move between distant patches via rare long distance
migration they will be associated with a much longer shared haplotype than
predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
results as a null model by which the contribution of long distribution
migrants to adaptation could be empirically judged.

We have also assumed that the alleles have a fitness effect in
heterozygotes both on and off the patch. However, 
 recessive alleles could also contribute to adaptation to patchy
 environments. \gc{should we cite Nachman eg here}.  Our
 simple branching process framework does not easily accomodate recessive
 alleles as it ignores the selection interaction of alleles, and so a
 different approach will be needed. At first sight it
 seems like recessive alleles may be able to contribute to adaptation
 via migration over longer distances, as the migrant alleles will be at low
 frequency between patches and so usually invisible to
 selection. However, this fact also means that they are unlikely to
 establish on the beneficial patch upon reaching it via migration. As
 such it is not clear what our intuition should be about the
 contribution of recessive alleles to adaptation via migration.

\gc{will write a couple more broader paragraphs}

The falling cost of sequencing over the next decade offers the
opportunity to investigate the genomics of parallel adaptation over a
large s
Over the coming decade population genomic data will be generated

\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}

if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
recall others' work.

patchy selection helps this happen; 
effective migration rate looks like what; 
strength of (positive?) selection suprisingly does not affect results (?)

standing variation will be important under these circumstances; 
migration rate (and so geographic structure) will not affect numbers of types if this is true; 
however migration always affcts spatial patterns.

recessive alleles will have different dynamics. Actually not really
obvious what will happen here, cline width will be much wider but
prob. of taking off will go as the square of the occupancy density.

discussion of applications.

discuss mixing of types; 
Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)

unequal density: Lenormand, gene flow and limits of\dots

small-scale fluctuations
when landscape is not made of discrete (circular) patches

>>>>

A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. \citep{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
On the lava flow the dark MC1R allele is at 86\% frequency,
while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
These numbers give us a sense of a plausible range of characteristic lengths.
Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
we can fit formula \eqref{eqn:eqfreq} to these frequencies.
Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
We also need estimate of $s_m$. 
Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
$1/\sqrt{30}$ are reasonable values.
\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
which has dispersal distance of 0.26km.  Shall we ask Nachman?}
\plr{He referred to \citet{french1968dispersal,allred1963range}.}

@@

This paper is an investigation into the basic question: 
What is the spatial resolution of convergent local adaptation?
In other words, what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
the dispersal distance divided by the square root of the selection
against the new allele between patches. It depends much more weakly on the
selective benefit within the patches or (perhaps suprisingly) the
population density. Although these two factors will determine the
time-scale over which adaptation will occur. This is contrast with models of panmictic populations
\citep{softsweepsI} and geographically spread populations adapting to homogeneous
selection pressures \citep{RalphCoop}, where the probability of
multiple adaptive alleles increases with the population
size/density. However, in all of these models the dependence on the beneficial selection
coefficient is absent or weak, due to the fact that selection both
aids establishment of new alleles and the spread of existing
alleles.

While weaker selection against alleles will make sharing of
adaptations between patches easier, 
it will also makes it harder to spot the sharing,
since the lucky, transiting alleles move slower and so carry a shorter
shared haplotype.
This issue is amplified by the fact that the length of haplotype
shared within patches decays over time, potentially making the
identification of shared adaptations to old selection pressures difficult.

% Observations of similar adaptations occuring within the same species in widely separated patches
% is strongly suggestive of parallel, independent adaptations.
% Such observations are very interesting 
% as they tell us about the degree of constraint faced by evolution --
% how many different methods are there typically available to a species
% with which to solve an evolutionary problem?

% It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
% \plr{fisher probably said this too?}
% and that spatial structure contributes to this \citep{ralph2010parallel}.

% This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
% Intermediate cases where habitat types are distributed in some more complex geometry
% -- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
% remains to be studied.

% To complement this theory, we need more empirical work,
% in which the parameters (migration rate, sensitivity to environmental factors, etc)
% are reasonably well-understood.
% However, learning about these aspects is difficult in the field,
% and it is still not understood how much of the intraspecific variation we observe
% is due to adaptation to local conditions, or even how to tell.
% Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?

@@

\paragraph{Dominance}
We have mostly swept the issue of dominance under the rug
by dealing with essentially haploid models,
and appealing to the fact that the dynamics we study occur where the mutation is rare,
and hence mostly present only in heterozygotes. 
Our results should hold as a good approximation to dominant and
partially dominant alleles.
If, however, the mutation is recessive, then it is essentially neutral where rare,
and so could have a much easier time spreading between patches.
This is counteracted, however, by the increased difficulty with which the mutation would establish
once arriving in a new patch. As
 such it is not clear what our intuition should be about the
 contribution of recessive alleles to adaptation via migration. 
Further work is needed to understand how recessive alleles may
contribute to convergent local adaptation.

\paragraph{Long distance migration}
We have also ignored the possibility of very long distance migration,
instead focusing local dispersal (as parameterized by a Gaussian
dispersal model). However, dispersal distributions can be very heavy
tailed, with a small fraction of individuals moving very long distances indeed, in
addition over long time-scales very rare chance events (mice sucked up in
hurricanes etc) could play a role in spreading migrant alleles. Such
tail events could greatly increase the probability of parallel
adaptation above that predicted by our model. Furthermore, if
adaptive alleles do move between distant patches via rare long distance
migration they will be associated with a much longer shared haplotype than
predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
results as a null model by which the contribution of long distribution
migrants to adaptation could be empirically judged.

The falling cost of population genomic sequencing means that we will soon have the
opportunity to study the effects of geography and ecology on
adaptation across many populations within a species. 
Our work suggests that even quite geographically close populations may
be forced to adapt convergently when migration is geographically limited. 
Thus, systems where populations have repeated to strong local
selection pressures limited may offer the opportunity to study
highly replicated convergent adaptation within a similar genetic
background.  
Such empricial work will also strongly inform our understanding of the role of gene flow in
keeping ecological similar populations evolving in concert \citep{sexton}.
Our results suggest that adaptation to shared environments is
certainly no guarentee of a shared genetic basis to adaptation, suggesting that
rapid adaptation to a share environment could potentially drive
speciation under some scenerios.

% \subsection{GRAHAM'S VERSION}

% We have developed a simple set of approximations to describe the rate
% at which organisms in a patchy environment adapt through novel
% mutations or migration from other nearby patches (eqn.
% \eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
% probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
% We have also determined the characteristic genomic length scales over which
% the various signals of hitchhiking should be visable
% (eqn. \eqref{eqn:haplotype_length} and ).

% These results show that the probability of parallel adaptation is
% independent of both the population density and the beneficial
% selection coefficient. This follows from the fact that these
% parameters both increase the rate of adaptation via both migration and
% mutation equally. Thus they are important for the time-scale over
% which adaptation occurs but not the mode of adaptation. This
% constrasts with models of convergent adaptation within panmictic
% populations \citep{} and geographically spread populations adapting to
% homogeneous populations \citep{RalphCoop}, where the number of
% adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
% coefficient is absent or weak, due to the fact that selection both
% aids establiishment of new alleles and the spread of existing
% alleles.

% The primary determinant of the probability of parallel adaptation is
% $\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
% allele frequency decay off the patch. We find that patches that are close together on
% this scale will adapt via migration, but there will be a sharp
% transition to convergent adaptation as we move to patches that are further apart.

% We have also assumed that the alleles have a fitness effect in
% heterozygotes both on and off the patch. However, 
%  recessive alleles could also contribute to adaptation to patchy
%  environments. \gc{should we cite Nachman eg here}.  Our
%  simple branching process framework does not easily accomodate recessive
%  alleles as it ignores the selection interaction of alleles, and so a
%  different approach will be needed. At first sight it
%  seems like recessive alleles may be able to contribute to adaptation
%  via migration over longer distances, as the migrant alleles will be at low
%  frequency between patches and so usually invisible to
%  selection. However, this fact also means that they are unlikely to
%  establish on the beneficial patch upon reaching it via migration. As
%  such it is not clear what our intuition should be about the
%  contribution of recessive alleles to adaptation via migration.

% \gc{will write a couple more broader paragraphs}

% The falling cost of sequencing over the next decade offers the
% opportunity to investigate the genomics of parallel adaptation over a
% large s
% Over the coming decade population genomic data will be generated

% \plr{Note that optimal $s_m$ for migration makes migration harder to spot.}

% if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
% recall others' work.

% patchy selection helps this happen; 
% effective migration rate looks like what; 
% strength of (positive?) selection suprisingly does not affect results (?)

% standing variation will be important under these circumstances; 
% migration rate (and so geographic structure) will not affect numbers of types if this is true; 
% however migration always affcts spatial patterns.

% recessive alleles will have different dynamics. Actually not really
% obvious what will happen here, cline width will be much wider but
% prob. of taking off will go as the square of the occupancy density.

% discussion of applications.

% discuss mixing of types; 
% Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)

% unequal density: Lenormand, gene flow and limits of\dots

% small-scale fluctuations
% when landscape is not made of discrete (circular) patches
